% $Id$

\documentclass[a4paper,12pt]{article}
\usepackage{amssymb}

\setlength{\parindent}{0mm}
\setlength{\parskip}{7.5mm}

\begin{document}

\title{Course 3BA7: Compiler Design I \\ Additional Lecture Notes \\ $14^{th}$ October 2004}

\maketitle

Syntax analyser ensures that the sequence of Lexical Tokens is
gramatically correct.

\section{Code Generation}

Expansion of sequence of atoms into object code ($\equiv$ machine
codeassembly language/core image memory image)


\section{Semantic Analysis}

Processing relating to the meaning of symbols.

Eg: semantics of an identifier.

	Includes it's type, it's bounds, if it is an array, etc...

Semantic processing includes:

	Static:

		Filling it's symbol table entry with a type of identifier.

	Dynamic:	

		Selecting correct class of operator for instruction when expanding
		atoms during code generation.

\begin{verbatim}

A + B
.
.
.
.
\end{verbatim}

ADD($\uparrow$ A, $\uparrow$ B, $\uparrow$ $T_{n}$)

% Diagram

Optimiser:

Optimisation code imporvisation techniques.


\section{Runtime Implementation}

Symbol table management

passes

\section{Finite State Machine}

i A finite state of imput symbols (inut alphabet)

ii A finite set of states, including a starting state.

iii A transition function  $\delta$ that assigns a new state to every
combination of state and input.

iv A set of accepting states

v A set of rejecting states.

\[ \delta(S_{OLD}, INPUT) \Rightarrow S_{NEW} \]

% Diagram

Eg: Parity Checker

Input: $0, 1$

States: $\{EVEN, ODD \}$

Starting State: $\{EVEN\}$

Transitions:	

$\delta(EVEN, 0) \Rightarrow EVEN $ \\
$\delta(EVEN, 1) \Rightarrow ODD $ \\
$\delta(ODD, 0) \Rightarrow ODD $ \\
$\delta(ODD, 0) \Rightarrow EVEN $ \\


Accepting States: ${ODD}$


Transition Table:

\begin{tabular}{|c|c|c|}
\hline
States/Input & $0$ 		& 	$1$ 	\\
\hline
$EVEN$		 & $EVEN$	& $ODD$	\\
\hline
$ODD$			 & $ODD$		& $EVEN$	\\
\hline
\end{tabular}

$1 = $ Accepting State \\

$0 = $ Rejecting State

Transition Diagram:

% Diagram

$01011$

% diagram

\begin{tabular}{|c|c|c|c|}
\hline
		 	& $0$ 	& $1$ 	& 		\\
\hline
$S$		& $S$		& $S$		& $0$	\\
\hline
\end{tabular}


\begin{tabular}{|c|c|c|c|}

\hline
		 	& $0$ 	& $1$ 	& 		\\
\hline
$S$		& $T$		& $T$		& $1$	\\
\hline
$T$		& $T$		& $T$		& $0$	\\
\hline

\end{tabular}


\section{Finite State Processing Machine:}

A finite state recognising machine with input alphabet extended by the
end marker symbol $\dashv$.


\section{Question}

Design a FSM to recognise input containing pairs of $1$s.


Input: $0$, $1$

States: ${NOONES, ONEONE}$

Starting State: ${NOONES}$

Accepting states: ${NOONES}$

Transition Table:

\begin{tabular}{|c|c|c|c|}
\hline
		 	& $0$ 			& $1$ 			& 			\\
\hline
$NOONES$	& $NOONES$		& $ONEONE$		& "YES"	\\
\hline
$ONEONE$	& $ERROR$		& $NOONES$		& "NO"	\\
\hline
$ERROR$	& $ERROR$		& $ERROR$		& "NO"	\\
\hline
\end{tabular}


\begin{tabular}{|c|c|c|c|}
\hline
		 	& $0$ 			& $1$ 			& $\dashv$		\\
\hline
$NOONES$	& $NOONES$		& $ONEONE$		& "YES" $0$		\\
\hline
$ONEONE$	& $ERROR$		& "YES" $1$		& "NO"			\\
\hline
\end{tabular}

\end{document}
