% $Id$

\documentclass[landscape,semcolor,a4]{seminar}
\usepackage{color}
\usepackage{pst-all}

\definecolor{darkblue}{rgb}{0,0,0.55}
\definecolor{conallblue}{rgb}{0.06,0.45,0.75}
\definecolor{white}{rgb}{1.00,1.00,1.00}

\newenvironment{stitle}{\LARGE\color{conallblue}}{}

\centerslidesfalse
\slideframe[\psset{linecolor=conallblue}]{scplain}

\pagestyle{empty}

\begin{document}

\sffamily
\pagecolor{darkblue}
\color{white}

\begin{slide}

\begin{stitle}
Presentation:
\end{stitle}

\vspace{15mm}

Document aesthetically well laid out. \\

\vspace{5mm}

Document language assumes a significant level of familiarity with broadband 
terminology and knowledge. eg DSL acronym (finally) explained on page 5 (of 8).

\end{slide}

\begin{slide}

\begin{stitle}
Objectives Met:
\end{stitle}

\vspace{15mm}

Provided criteria met, except:

\vspace{5mm}

Lack of info on Digiweb's other products, such as Satellite based
internet connectivity. \\

\vspace{5mm}

Lack of reference to competitors other than Eircom or competitor's
products or the differences (advantages/disadvantages) between DSL and
wireless broadband products. 

\end{slide}

\begin{slide}

\begin{stitle}
Improvements:
\end{stitle}

\vspace{15mm}

Consider the document's audience. Assume the reader has \textbf{no} knowledge 
about the broadband services sector and is researching the major
providers in Ireland. \\

\vspace{5mm}

Briefly overview of the low and high end packages key features
(such as Download/Upload speeds, contention ratio and quota) to allow
effective comparison with competitor products. Minor package details 
such as email features and the cost of each package at the time of 
writing are irrelevant and subject to change.

\end{slide}

\end{document}
