% $Id$

\documentclass[a4paper,12pt]{article}
\usepackage{amssymb}

\setlength{\parindent}{0mm}
\setlength{\parskip}{7.5mm}

\begin{document}

\title{Course 3BA4: Computer Architecture II \\ Lecture Notes \\ $8^{th}$ November 2004}

\maketitle

\section*{Comparison of DLX Branch Strategies}

An alternative strategy is to use delayed branches (like RISC-I and
Sparc)

Assume:

\indent $14\%$ Branch Instructions

\indent $65\%$ of branches change PC

\indent $50\%$ probability of filling a branch delay slot with a useful
instruction.


\begin{tabular}{l|l|l}
Method				&	Branch Penalty	&	Effective CPI				\\
\hline
stall pipeline	3	&	$3$				&	$0.86 + 0.14*4 = 1.42$	\\
\hline
stall pipeline	1	&	$1$				&	$0.86 + 0.14*2 = 1.14$	\\
\hline
predict not taken	&	$1$				&	$0.86 + 0.14*(0.35 + 2 * 0.65) = 1.09$	\\
\hline
delayed branch		&	$0.5$				&	$0.86 + 0.14*1.5 = 1.07$ \\
\end{tabular}

\section*{Branch Prediction}

Need to resolve branches in IF phase

Use a branch target buffer (cache)

Contains the PCs of the branch instructions executed.


\begin{tabular}{|l|l|l|}
\hline
PC				&		Predicted PC		&	Branch Prediction State	\\
				&								&									\\	
				&								&									\\	
				&								&									\\	
				&								&									\\	
				&								&									\\	
				&								&									\\	
\hline

				&	Output Predicted PC if Match Found	&									\\	
\end{tabular}


During IF stage, match Pc against all PCs in Branch Target Buffer

If match is found, use predicted PC in IF phase of next instruction

If branch predicted, pipeline \textbf{not} stalled.

If branch incorrectly predicted, must \emph{"abort"} fetched instruction
and fetch correct one, pipeline stalled.

Must update branch target buffer if a \emph{"new"} branch fetched or
prediction changes.

\section*{Two Bit Branch Prediction}

Consider a loop which is executed 10 times

\begin{eqnarray*}
\verb!r1! & \leftarrow	& \verb!10! \\
		  . & 	& \\
		  . & 	& \\
\verb!r1! & \leftarrow	& \verb!r1 - 1! \\
			 & \verb!bnez!	& \verb!r1,L1! \\
\end{eqnarray*}

Assume a branch is predicted to branch the same way as it did the last
time.

Branch taken $9$ times and then not taken.

But \textbf{next} loop is executed first branch will be predicted
incorrectly.

\indent $\Rightarrow 80\%$ prediction accuracy

A bwo bit scheme changes prediction \textbf{only} when predicted
incorrectly twice in succession.

\indent $\Rightarrow 90\%$ prediction accuracy (in this example)


\end{document}
