% $Id$

I wish to thank my supervisor, Stephen Barrett who not only
agreed to take on this project, he let me run with my ideas, and 
provided wisdom which helped me though out the whole period.

\vspace{5mm}

Grateful thanks is owed to: (in no particular order) David Malone, 
Ian Dowse, Gary Coady, Ian O'Connell, Aidan Kehoe and 
Conor McDermottroe, all of who helped spot logical and algorthimic 
mistakes, as well as always being willing to answer C related 
questions.


I also want to acknowledge Adrian Colley, Paul McGaley, Cian Synnott, 
Colm MacC\'{a}rthaigh and Colin Whittaker for providing feedback and 
ideas during the planning stage.


Very special thanks is due to Niall Richard Murphy, who not only was 
willing to answer C related queries, but was generous enough to provide 
access to a remote system as a replacement build environment after 
unforeseen difficulties.

\vspace{5mm}

I wish to thank Ciara Brennan, Sharon Murphy and N\'{o}ir\'{i}n Plunkett
for proof reading this report. 


Special thanks are due to Edsko de Vries for help with \textrm{\LaTeX{}} 
and \textrm{\Xy-pic} queries during the preparation of this report.

\vspace{5mm}

I want to thank John Tobin for providing support and advise, in 
addition creating the core system model and original code-base.

\vspace{5mm}

Finally, I would like to thank my father, who once upon a time, 
asked his 5 year old son to assist him to help reconfigure jumpers on
a motherboard. He didn't realise at the time, what he had started...
