% $Id$

\subsection{A New Perspective}

Many of the specified features for this project have been interests of
mine for quite a while. In particular, I regularly use numerous flavours
of BSD, Unix and Linux, Unicode, GnuPG and IPv6 networking. However, this 
was the first time I developed software in any of these areas.


Over the course of this project, I learned that my knowledge and general
usage of these technologies is quite different to programming in these
areas. Until now I never "\emph{got my hands dirty}". Now that I have, I
have a greater understanding of these technologies and a greater
appreciation for the complexities and issues associated with developing
applications in these areas.

\subsection{Programming Experience}

I was partially treating this project as a reason to learn and practice
C programming. Prior to undertaking this project, my C programming ability
was that of an intermediate, but inexperienced, programmer; I could read 
and understand basic and common C code, but I had limited experience
writing significant applications, algorithms or data structures through
C.


This inexperience was apparent as development commenced. After a short
period of time, as I began to integrate various libraries together, I
became more comfortable working in C. Towards the end of the
development process, I was using more sophisticated C controls and
approaches that had once confused me. As a result of this project, I am
more familiar and comfortable working in the C programming language.


\subsection{Project Management}


Managing a project is not an easy task. Not only does a manager need to
contribute to the team, but they must also organise the team, delegate
tasks and set deadlines in addition to envisioning the overall picture.


I am comfortable admitting that I was not managing this project properly
at the start. My primary focus was on researching various technologies
and tools to aid the development and documenting stages of the project.


At this time, deadlines were set with the belief that there was plenty
of time left. This was an assumption that should not have been made.
Various other commitments, primarily academic and professional, impacted
on this available time.


Once I became pro-active and began managing various academic, personal,
professional and project tasks, my project management improved. Instead
of focusing on simply research and useful tools, I was developing the
core, essential components, while concurrently researching various other
technologies and setting practical, realistic deadlines for milestones.


As a result, I became more efficient and productivity was more and more
noticeable. Had these project and time management techniques been
applied from the start of the project, I am certain that more, if not
all, of the specified features would have been implemented.

\subsection{Working Status}

Finally, I am happy with the concluding status of this project. While I
did not achieve more than one of the features I set out to implement, I
delivered a working, albeit crude, distributed messaging system. In
addition, I have produced a well thought out road-map and, if 
development were to continue, I am confident that the specified features
would be easily achieved and expanded upon. 

