% $Id: /college/fyp/2.0/doc/objective.tex 2490 2006-03-19T22:58:59.629349Z conall  $

The client should be intuitive for a novice to use, but versatile for a
power user.


The client should transmit the message to the server on the destination
host. The server then attempts to delivery of the message to the
recipient and transmit status details to the sender via the client.


The system should support IPv4 and IPv6 sockets to communicate between
client and server. 


The system should be as portable as possible, compiling and working on 
various Unix and Unix related systems (including commercial Unix
implementations like Solaris and MacOS X, Unix 
implementations such as the primary BSD flavours, and open source Unix
related systems such as various Linux distributions) 


Messages between client and server should be cryptographically secure by
default. Due to the additional security concerns using SSL connections
and corresponding X.509 certificates to encrypt messages in transit
between servers on separate nodes. Public-private keys should
be used and the encryption should be from end user to end user.
OpenPGP should be the encryption model used.


The client should support unicode input from the user. The server should 
attempt delivery in Unicode, if the terminal of the recipient is Unicode 
capable. 


Messages should be transmitted in unicode. Non unicode input should be
encoded into unicode prior to encryption and transmission. Messages
should be decoded from unicode after transmission and decryption, if
the terminal of the recipient does not support unicode. 


The server should be able to join IP multicast group, to allow multiple
servers to receive the same message transmitted by a client to an IP
multicast group. 
