\documentclass[a4paper,12pt]{article}
\usepackage{amssymb}

\begin{document}

\title{Course 3BA2: Artificial Intellegence \\ Additional Lecture Notes \\ $18^{th}$ October 2004}

\maketitle

\section*{Prolog Syntax}

Everything is a term. \\

\\

Terms are simple or complex ("Structured").

\subsection*{Simple}

"atom" - stands for something. Eg: mike, chair \\

\\

\subsubsection*{Syntax:} 

start with a lower case letter and followed by zero or more 
alphanumeric characters or underscores.


A special character followed by zero or more special characters: $ \&, \#. /. \*, \$ $


Any string of characters enclosed in single quotes. Eg: 'The w cat', 'foo.txt', 'can''t do it'

\subsubsection*{Numbers:}

Integers \\

Floats (Implementation dependant)

\subsubsection{Variables:}

Anonymous Variable: $\_$ \\

member(x,[_|R]). \\

\subsubsection{Named Variable:}

Capitol Letter followed by zero or more aplhanumeric characters \\

or \\

An underscrore followed by one or more alphanumerics or underscores. \\


Scope of variable in a clause is the clause itself if named or no scope if anonymous.

\subsection{Complex:}

has a functor, which has a name (some syntax rules as an atom) and has an arity of $\gt 0$. \\

composs_points(north,south,east,west). \\

north - Prolog term \\

tree(_,R). \\

temp(L1,R). \\

foo(X). \\

bar(X,R) $ : $ foo(X),foo(X). \\

$ : $ - Infix operator \\

inter operator - binary - $x + 1$ \\

Postfix operator - $24$ \\

Prefix operator - ??? \\

Prefix operator - ??? \\

Prefix operator - ??? \\

\\

:- (bar(X,Y),\left(foo(X),foo(Y)\right)

\subsection{Lists in Prolog}

A list has 2 arguements, the 'head' and the 'tail'/'head' and 'body'/'car' and 'cdr' (LISP method) \\

.('head','body').

[a,b,c]

.(a,.(b,.(c,[]))).

[] - Emplty list atom

Prolog is a logic programming language.

Clauses are axioms.

When running a prolog program, you're asking for a proof that the query you make is true.


member(X,[X|_]).

member(X,[_|R]) :- member(X,R).


eg: query:


member(a,[61,4,9,a,a(b),X]).


Prolog makes Logical Interfaces to prove this goal.


In C:


% Diagram


In prolog:


member(a,[61,a]).
   
member(x,[X|_]).


attempt to unify each arguement with the goal with tit's corresponding arguement in the head.

	- if the arguements are identical
	
	- if one is an unbound variable and the other is not, then the unbound variable is 
	- made equal to the other term (it is instantiated). The two terms are then identical
	  In fact, they are the same term from then on.

	- If both are unbound variables, they are unified such that they are identical there after,
	  the same unbound variable.

	- Otherwise, unification fails and resolution fails.

\end{document}
