 % $Id: /college/trunk/3ba7/I/Notes/20050111.tex 293 2005-04-26T19:21:47.134940Z conall  $

\documentclass[a4paper,12pt]{article}
\usepackage[usenames]{color}
\usepackage[arc,poly,all]{xy}

\setlength{\parindent}{0mm}
\setlength{\parskip}{7.5mm}


\begin{document}

\section{Client Sending a Message to Server}

\xy<1cm,0cm>:
(10,0) *=(7.5,1)!UC\txt{Message} *\frm{-} ;
(10,-3) **\dir{-} *\dir{>},
(10,-3)*=(7.5,1)!UC\txt{Encode into Unicode} *\frm{=} ; 
(10,-6) **\dir{-} *\dir{>},
(10,-6)*=(7.5,1)!UC\txt{Encrypt Message} *\frm{-} ; 
(10,-9) **\dir{-} *\dir{>},
(10,-9)*=(7.5,1)!UC\txt{Passed to $Server_{Sender}$} *\frm{-} ; 
(10,-11) **\dir{=} *\dir{>},
\endxy

\begin{enumerate}

\item User types messages, then hits send.

\item Message is (optionally) translated into Unicode. If message is
already in Unicode, no translation is performed.

\item Message is encrypted before transport across the network

\item Message is placed in the outgoing queue of the local server

\end{enumerate}

\section{Sending a Message between Servers}

\xy<1cm,0cm>:
(10,0) *=(7.5,1)!UC\txt{Send from $Server_{Sender}$} *\frm{-} ;
(10,-3) **\dir{-} *\dir{>},
(10,-3)*=(7.5,1)!UC\txt{Receive at $Server_{Receive}$} *\frm{-} , 
(7.5,-4) ; (7.5,-6)**\dir{-} *\dir{>},
(12.5,-6) ; (12.5,-4)**\dir{--} *\dir{>},
(10,-6)*=(7.5,1)!UC\txt{Poll for User} *\frm{-} ; 
(10,-9) **\dir{=} *\dir{>},
\endxy

\begin{enumerate}

\item Message is queued in $Server_{Sender}$.

\item $Server_{Sender}$ passes message to $Server_{Receiver}$.

\item $Server_{Receiver}$ polls for recipient user. 

\subitem If user is available, attempt to delivery messege. 

\subitem If user is not available, pass message to local SMTP queue. 

\subitem If user does not exist, destroy message and report error to $Server_{Sender}$.

\end{enumerate}

\section{User Receiving a Message from Server}

\xy<1cm,0cm>:
(10,2) ; (10,0) **\dir{=} *\dir{>},
(10,0) *=(7.5,1)!UC\txt{Message Received by $Server_{Reciece}$} *\frm{-} ;
(10,-3) **\dir{-} *\dir{>},
(10,-3)*=(7.5,1)!UC\txt{User is Available} *\frm{-} ; 
(10,-6) **\dir{-} *\dir{>},
(10,-6)*=(7.5,1)!UC\txt{Decrypt Message} *\frm{-} ; 
(10,-9) **\dir{-} *\dir{>},
(10,-9)*=(7.5,1)!UC\txt{Decode from Unicode} *\frm{=} ; 
(10,-12) **\dir{-} *\dir{>},
(10,-12)*=(7.5,1)!UC\txt{Print Message to User} *\frm{-} ; 
\endxy

\begin{enumerate}

\item Message is received by $Server_{Receive}$.

\item User appears to be available. Attempting to deliver messege.

\item Message is decrypted by recipient user. 

\item Message is optionally translated from Unicode, unless user has
declared they can accept Unicode. 

\item Message is printed to user's terminal

\end{enumerate}

\end{document}
