\documentclass[a4paper,12pt]{article}
\usepackage[all]{xy}

\begin{document}

\title{2BA6 \\ Computer and Society \\ Michaelmas Term 2003 \\ Lecture
Notes}

\author{}%Conall O'Brien}

\date{}

\maketitle

\section*{Week 1 - 08/10/2003}

\noindent 1 Sun Day \\
\noindent 2 Moon Day \\
\noindent 3 Mars Day\\
\noindent 4 Mercury Day\\
\noindent 5 Jupiter Day\\
\noindent 6 Venus Day\\
\noindent 7 Saturn Day\\


\noindent Computer Science and Society

\vspace{10mm}

\noindent History and Philosophy of Science

\vspace{10mm}

\noindent Story told from the perspective of the telling

\vspace{10mm}

\noindent "The Structure of Scientific Revolutions" - Thomas S. Kuhn

\vspace{10mm}

\subsection*{Questions and Answers}

\noindent Who made the world?

\vspace{10mm}

\noindent How did the world  come into being?

\vspace{10mm}

\noindent How did time start?

\vspace{10mm}

\noindent Le Sens Pratique de Philosphant (the Practival Sense of Philosophy)

\vspace{10mm}

\noindent The Act of Asking the Right Questions

\vspace{15mm}

\noindent Q. Does it exist or is it only in my "mind"?

\section*{Week 2 - 15/10/2003}

No Lecture

\section*{Week 3 - 22/10/2003}

\noindent To Do

\section*{Week 4 - 29/10/2003}

\noindent One page essay in HTML format on a "Scientific Revolution related topic
from the course textbook". - Due next Wednesday.

\vspace{10mm}

\noindent 1984 - Orwell - Newspeak

\vspace{10mm}

\noindent Ponte - find out about %something

\vspace{5mm}

\noindent Bruno

\vspace{5mm}

\noindent Vico - History

\vspace{5mm}

\noindent Joyce

\vspace{5mm}

\noindent Beckett

\vspace{10mm}

\noindent Tin Drum - Gunter Grass - Crabwalk

\section*{Week 5 - 05/11/2003}

\noindent "Oryx and Crake" - Margaret Atwood (Canadian)

\vspace{10mm}

\noindent "Snow Crash" - Stephenson

\vspace{10mm}

\noindent Kuhn's Trigger - 

\vspace{10mm}

\noindent Aristotle - "Laws of Motion" \\
\noindent Thomas S. Kuhn - "The Copernican Revolution" (1957) - page 75 \\
\noindent Thomas S. Kuhn - "The Essential Tension" (1977)

\vspace{10mm}

\noindent Computer Science Revolution

\vspace{10mm}

\noindent Socrates - learn how to ask questions

\vspace{5mm}

\noindent Plato

\vspace{5mm}

\noindent Aristotle

\vspace{5mm}

\noindent Theopharstus

\section*{Week 6 - 12/11/2003}

\noindent Method - way to find knowledge

\vspace{10mm}

\noindent Socratic - asking questions - manner or method of framing

\vspace{10mm}

\noindent Bizzme language - easier to remember eg UNIX. \\

\noindent - Airline Pilots \\
\noindent - UNIX Programs \\
\noindent - Priests (Religion)

\vspace{10mm}

\noindent "My goal is to pervert you for life"

\vspace{10mm}

\noindent Astronomy: Keplen - Ellipse, Observations, Measurements

% Ellipse

\vspace{10mm}

\noindent Simple Harmonic Motion

\section{Week 7 - 19/11/2003}

\noindent 1. Exam - 3rd December - Open Book

\vspace{10mm}

\noindent Covered Material of current term: \\

\noindent - Lectures
\noindent - Kuhn

\vspace{10mm}

\noindent 2. Group Project (Max 5 people) \\
\noindent Hilary Term \\
\noindent Subject Matter

\vspace{10mm}

\noindent Information $\to$ "Google News 4500".

\vspace{10mm}

\noindent 3.

\noindent What is the point of this course - 2BA6? \\
\noindent "Teaching $\longleftrightarrow$ Research"

\vspace{10mm}

\noindent Exam:

\vspace{10mm}

\noindent "What is this thing called Computer Science?"

\vspace{10mm}

\noindent Commnt on History and Philosphy of Science. \\

Danta ... Bruno ... Nico ... Joyce

\vspace{10mm}

\noindent Comment on Crabwalk (In Krebsqang).

\vspace{10mm}

\noindent Comment on the days of the week with respect to Kuhn's
"Structure of Scientific Revolutions".


\section*{Week 8 - 26/11/2003}

Weapons of Mass Deception - Sheldon Rampton, John C. Stauber

\vspace{10mm}

\noindent - Reuters \\
\noindent - Associated Press

\vspace{10mm}

\noindent Information

\vspace{5mm}

\noindent INFO - MATION \\
\noindent INFORM ATION

\vspace{10mm}

\noindent History -- Latinised

\vspace{5mm}

\noindent Herstory

\vspace{10mm}

\noindent Author(ity)

\vspace{10mm}

\noindent Information $\to$ Knowledge $\to$ Wisdom

\end{document}
