% $Id$

\documentclass[a4paper,12pt]{article}

%\usepackage{eurosym}

\setlength{\parindent}{0mm}
\setlength{\parskip}{7.5mm}

\newcommand{\euro}{\kern.17em{}C\kern-.95em\raise.19ex\hbox{\it
=\/}\kern.17em{}}

\begin{document}

\title{Semtex Solutions Ltd \\ Business Plan}

\author{Vivian Hunt \\ Stephen Kealy \\ Conall O'Brien}

\maketitle

\newpage

%\index

\section{Introduction}

This business plan is the result of a group effort from the three
management partners of Semtex Solutions Ltd. It was written to help
document the core strategies involved in the development, production and
marketing of our product. It will also act as a tool for the managers of
Semtex Solutions to moniter the performance of the business. In the
later stages of the businesses growth, this document will serve to aid
the management team in evaluating the businesses effectiveness with
regards to competition. Also, our business plan will be used to
ascertain if our business targets are being met, and that we adhere to
our mission statement at all times.

Our range of products and services is quite a large undertaking for a
small group of individuals to raise capital for. To this end, the
business start up stage will rely on receiving funding from potential
business investors such as venture capitalist and IT related
organisations or technology related firms. This business plan documents
how and where equity funding will potentially come from. Setting up this
business will require some aid from government bodies. Included in this
plan is a full detailed documentation of who and where we intend to get
this aid from. 

Background

\begin{tabular}{ll}
Business Name	& 	Semtex Solutions Ltd.	\\
Address			&	Trinity College Dublin, \\
					&	Dublin	2					\\
					&	Ireland						\\
					&									\\					
Telephone 		&	(01) 687453					\\
Facsimile: 		&	(01) 453281					\\
E-Mail:			&	info@SemtexSolutions.ie \\
					&									\\					
Status:			&	Partnership based Limited Company \\
Registered for &	VAT, PAYE/PRSI, Corporation Tax, Income Tax \\
Formed as		&	Start-Up						\\
					&									\\					
Business Start Date	& 	15th October 2004	\\
\end{tabular}

\section{Management}

\subsection{Vivian Hunt}

Position: Chief Executive Officer and Financial Controller, Chief
Programmer.


Vivian, is the youngest CEO of a Ltd company in Ireland. His career
began in Mayo producing peat based fuels to the rest of the west. Vivian
then moved into the hardware distribution franchise, currently known as
Arrow Hardware. His first experience in the media industry came when he
was offered the co-management position of a video rental store. Here, he
acquired in dept knowledge of the industry and is now an expert in this
field. Vivian then thought it prudent to study the dynamic field of
computer science in Trinity College Dublin where he is now in his third
year. It was in this fine college that he headed the new company, Semtex
Solutions.

\subsection{Stephen Kealy}

Executive Director and Marketing Manager, Database Administrator.


Stephen has had a very colourful and vibrant career. After completing
his schooling in 1990 Stephen attended the University of Australia where
he completed a degree in Commerce. In 1994 He joined National Australia
Bank and in 1995 he became the youngest senior accountant with National
Australia Bank. In 1998 Stephen joined the United Nations to oversee the
financial control of oil sales to all of Africa. In 2001 Stephen Arrived
in Dublin to discover his roots. In doing so he discovered a love of TV
and Computers which led him to Trinity College Dublin where he is
currently a third year Computer Science Student.


\subsection{Conall O'Brien}

Non Executive Director and Chief Sales Officer, Networks Administrator. 


Following his graduation from Blackrock College in 1995 and graduating
Trinity College with a BA in Computer Science, Conall has worked in the
Irish IT sector for companies including Trinity College, Critical Path
and the INEX where he was actively involved in the design and
implementation of various networks and rising into management positions.

\section{Product/Service}

Our business can be classed as a provider of high quality media content.
It uses the relatively new technology that is broadband, to deliver all
available media content from television stations, television programmes
and movies to the registered user. TeleBand is the name of our product.
The name encapsulates both the product and the service being provided.
The product .TeleBand. in its physical form is a digital box that
connects directly to a highspeed internet connection and a television. A
universal remote control is included for convenience. TeleBand also
comes as a software package. The software package allows the user to use
the TeleBand service from a laptop or a Personal Computer with a
broadband connection. The software package will allow the user to view
media files of all format from the most popular like Mp3 and DivX.


The core aspect of the business is the service it provides. The service
can be split into different categories according to the type of content
provided. These categories are; On Demand Video, On Demand Television
Programmes, TeleBand Television and TeleBand Digital. 


On Demand Video is the option that lets the TeleBand subscriber view the
extensive list of films available. The user can then download and view
the movie they wish to see. The users account is charged for the movie.


On Demand Television Programmes is another option that lets the TeleBand
client chose what television they want to watch. This is different to
the Sky Plus service in that the user can view what they want, when they
want it. A long and extensive list of television programmes is available
for view at the users discretion. This alleviates the need to record
your favourite television programmes as they are available at a pay per
view and on demand basis.


TeleBand Television is the service that allows TeleBand clients to view
the existing television stations such as RTE, TV3 and the BBC.s on their
computers or through the TeleBand box. This service is free.


TeleBand Digital is another service that allows users to subscribe to a
digital television service provider such as NTL and view digital
television through their TeleBand Box or TeleBand Media Environment
software.


\subsection{Mission Statement}

Our mission is to establish a stronghold in the online entertainment
market. We intend to be the largest legitimate video media content
provider on the world wide web. We will do this by;

\begin{enumerate}

\item	Providing a product that is easy to use and navigate

\item	Providing a constant reliable service that delivers 
		high quality all the way

\item	Giving the user what he/she wants.
\end{enumerate}

\subsection{Trends in Industry}

The changing trends in how the internet is used and the increasing
numbers of people online are the basis for this products potential. More
people are getting access to the internet every day and its not just
connection that is increasing. Ireland was one of the last European
countries to get broadband. In the few years that broadband has been
here, broadband providers are appearing everywhere. We are now seeing an
upsurge in the numbers of domestic households that have broadband
internet access. Today, the common domestic broadband service provides
information transfer at a rate of two megabytes per second. This is more
than adequate to support a product like TeleBand without affecting
everyday internet usage.


More people are buying online. The year 2004 has seen record numbers
going to the internet to do their shopping. Consumers are now looking to
the internet to buy their groceries, Christmas and birthday presents,
music, concert tickets and other products. This trend is still
increasing today. There is every reason to believe that these trends
will continue to increase and that the range of products and services
available online is set to expand.


Entertainment is one of the main ways people use the internet. The type
of entertainment we are concerned with is video and music. The rate of
new media formats appearing is a trend in itself. Usage of these formats
according to recent surveys are on the increase. There are several
companies that provide different entertainment services such as music
download in Mp3 format. These sites and software packages litter the
web. However, as of right now, there are only a few video rental and
movie download outlets online, let alone television on demand. 


These three trends, we believe, will pave the way for an innovative
product like TeleBand to fill this niche in the market. 


\subsection{Targets}

In the present moment, our targets are to develop our product and
service fully and to expand our servers television and video content.
Developing our product will involve having the digital box and remote
designed and produced. Developing our service will involve purchasing
fast servers to provide the service, programming the service and
administration. Expanding our service's content will involve getting
contracts with television and movie production companies so we can
provide their programmes and films to the user.

\section{Marketing}

\section{Sales}

\section{Operations}

\section{Financial}

Since this is a startup company offering a product and service currently
unrivalled, the market place doesn't exist. As a result, a conservative
spending approach is taken initially in order to gauge public interest 
in the product...

\subsection{Overall Startup Budget}

Start up capital of $\euro 1,000,000$

\subsubsection{Acquired through grants from}

\begin{tabular}{lccr}
Enterprise Ireland								&	&	$\euro 25,000$ \\
banks loads											&	&	$\euro 75,000$ \\
Technology Venture Capitalists				&	&	$\euro 800,000$\\
Management Team									&	&	$\euro 100,000$
\end{tabular}

\subsubsection{Operational Costs}

\begin{tabular}{lccr}
Research and Development						&	&	$\euro 150,000$ \\
Product Development     						&	&	$\euro 250,000$ \\
Management Team         						&	&	$\euro 400,000$ \\
Reserve Fund   						         &	&	$\euro 200,000$
\end{tabular}

\subsection{Initial Operating Budget}

Start Operational Budget of $\euro 400,000$

\subsection{Operational Costs}

\begin{tabular}{lccr}
Salaries      										&	&	$\euro 200,000$ \\
Overheard      									&	&	$\euro 100,000$ \\
Infrastructure 									&	&	$\euro 100,000$
\end{tabular}

\end{document}
